---
title: 'Activity: Bird Species'
author: "Your Names Here"
date: 'DUE DATE '
output:
  html_document:
    df_print: paged
---


# Set Up

```{r}

# load packages
library(tidyverse)


# Load the BabyNames data into RStudio environment (from dcData package)
data("OrdwayBirds", package = "dcData")

# Load the `OrdwayBirds` data set into your RStudio environment from the `dcData` package

OrdwayBirds <-
  OrdwayBirds %>%
  select( SpeciesName, Month, Day ) %>%
  mutate( Month = as.numeric(as.character(Month)), 
          Day = as.numeric(as.character(Day)))


# Show a few commands used to inspect the data set a few different ways

head(OrdwayBirds)
str(OrdwayBirds)
summary(OrdwayBirds)

```

# Activity

When you're finished with this activity, you will have a graph that shows what time of year various species appear at the Katherine Ordway Natural History Study Area in Inver Grove Heights, MN.

### Step 0

Before we begin, the book gives us some instructions to select a few key variables and clean up the date formatting.  Use the commands given in the book, but be sure to study the functions used so you know what is happening.  In this case, we select `SpeciesName`, `Month`, and `Day`. Then a `mutate()` command converts each variable to character, and then converts the resulting "character string" to a number.

```{r}

OrdwayBirds_cleaned <- OrdwayBirds %>%
  select(SpeciesName, Month, Day) %>%
  mutate(Month = as.numeric(as.character(Month)), 
         Day = as.numeric(as.character(Day)))


head(OrdwayBirds_cleaned)





```



### Step 1








```{r}
# Unique species in original data (including mis-spellings)
unique_species_original <- unique(OrdwayBirds$SpeciesName)

# Unique species in the clean list
unique_species_cleaned <- unique(OrdwayBirds_cleaned$SpeciesName)


```


### Step 2
  



```{r}

# Load the OrdwaySpeciesNames table
data("OrdwayBirds", package = "dcData")

# Correct the mis-spellings in SpeciesNames using inner_join
corrected <- OrdwayBirds_cleaned %>%
  inner_join(OrdwaySpeciesNames) %>% 
  select(Species = SpeciesNameCleaned, Month, Day) %>%
  na.omit() %>%  # Clean up the missing values
  distinct()  # Remove duplicate rows resulting from many-to-many relationship

# Inspect the corrected data
head(corrected)



```


### Step 3

We want to identify the top 5-6 bird species sighted at the Katherine Ordway Natural History Study Area.

```{r}
library(dplyr)
library(tidyverse)
# Count sightings by species in descending order

species_counts <- joined_data %>% count(SpeciesName, sort = TRUE)
# Display the top 10 species
top_species <- head(species_counts, 10)



# define a sightings threshold; produce a table with only birds belonging to major species
sightings_threshold <- 10
major_species <- top_species %>%
  filter(n >= sightings_threshold)



```


### Step 4

We now want to produce a month-by-month count for each of the major species, and then construct a chart to tell the story of what time of year the various species appear.

```{r}
library(dplyr)
# Prepare glyph-ready data table called `ByMonth`
ByMonth <- joined_data %>%
  group_by(SpeciesName, Month = Month.x) %>%
  count()



ByMonth %>% 
  ggplot(aes(x = Month, y = n, color = SpeciesName)) + 
  geom_line()

ByMonth %>% ggplot(aes(x=Month,y=n ,fill=SpeciesName)) +
  geom_bar(stat='identity',position='stack', width=.9) 

```

Now use the ggplot template below to generate the desired graph. (Remark you can also use esquisser or mplot to create a line chart instead of the barchart)





According to the graph, we can answer the questions posed:

##### 1. Which species are present year-round?  
1 American Goldfinch     
2 Black-capped Chickadee 
3 Blue Jay               
4 White-breasted Nuthatch


##### 2. Which species are migratory, that is, primarily present in one or two seasons?
1 ""                   
 2 "-lost-"             
 3 "-missing-"          
 4 "13:00:00"           
 5 "Acadian Flycatcher" 
 6 "American Golf Finch"
 7 "American Redfinch"  
 8 "American Redstart"  
 9 "American Robin"     
10 "Arkansas Kingbird"



##### 3. What is the peak month for each major species? (bullet list)
peak_month
  SpeciesName           Month
   <chr>                 <dbl>
 1 ""                       NA
 2 "-lost-"                 10
 3 "-missing-"              11
 4 "13:00:00"                2
 5 "Acadian Flycatcher"      8
 6 "American Gold Finch"     5
 7 "American Goldfinch"     10
 8 "American Golf Finch"     3
 9 "American Redfinch"       5
10 "American Redstart"       5



##### 4. Which major species are seen in good numbers for at least 6 months of the year
1 American Goldfinch     
2 Black-capped Chickadee 
3 Blue Jay               
4 White-breasted Nuthatch             



```{r}

# Filter species present year-round
species_year_round <- ByMonth %>%
  group_by(SpeciesName) %>%
  filter(n_distinct(Month) == 12) %>%
  distinct(SpeciesName)

# Filter species primarily present in one or two seasons
species_present <- ByMonth %>%
  group_by(SpeciesName) %>%
  summarise(NumSeasons = n_distinct(Month)) %>%
  filter(NumSeasons <= 2) %>%
  distinct(SpeciesName)


# Find the peak month for each major species
peak_month <- ByMonth %>%
  group_by(SpeciesName) %>%
  filter(n == max(n)) %>%
  select(SpeciesName, Month) %>%
  distinct()




# Filter for major species with sightings in at least 6 months
major_species <- OrdwayBirds_cleaned %>%
  group_by(SpeciesName) %>%
  summarise(NumMonths = n_distinct(Month)) %>%
  filter(NumMonths >= 6)




```

